---
title: "Portfolio 4: Demographic and Pre-Screen Info"
Date: "4/22/2025"
output: html_document
---

```{r random, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As per journal guidelines, I had to report demographic information for my submission. Additionally, I reported some brief descriptive information on the major life events experienced by my sample, which was pertinent to the research question. For this portfolio piece, I will be recreating these preliminary analyses with my newly minted .rds files. 

```{r setup}

library(tidyverse)
library(readr)
library(janitor)

demographic_data <- readRDS("demographic_data.rds")
life_event_data <- readRDS("life_event_data.rds")

```

## Demographics

```{r cleanup}

demographic_data <- demographic_data %>% clean_names()

names(demographic_data)

#had to clean these variables out
demographic_data <- demographic_data %>%
  select(-matches("^x(7|8|9|10|11|12|13)$"))

```

```{r descriptives_dems}

# Age
demographic_data %>%
  summarise(
    mean_Age = mean(age, na.rm = TRUE),
    sd_Age = sd(age, na.rm = TRUE),
    min_Age = min(age, na.rm = TRUE),
    max_Age = max(age, na.rm = TRUE)
  )

# Sex
demographic_data %>%
  count(sex) %>%
  mutate(prop = round(n / sum(n), 3))

# Cleaner categorical tables
demographic_data %>% tabyl("ethnicity_simplified")
demographic_data %>% tabyl("student_status")
demographic_data %>% tabyl("employment_status")

```


This is great, and mimmicks what I found and reported on excel. Now lets move onto the life events pre-screen data I have. 

## Pre-Screen Life-EVents

```{r pre_screen}

names(life_event_data)

```
```{r cleaning}

# The variable names are hideous so lets fix that and remove some of the unnecessary ones

library(dplyr)

life_event_data <- life_event_data %>%
  select(-c(18:22))

life_event_data <- life_event_data %>%
  rename(
    separated_partner = Separatedfromspouseorlongtermpartner,
    injury_self = Seriouspersonalinjuryorillnesstoself,
    injury_family = Seriouspersonalinjuryorillnesstocloserelativefamilymember,
    death_spouse_child = Deathofspouseorchild,
    death_relative = Deathofothercloserelativefamilymembere.g.parentorsibling,
    death_friend = Deathofaclosefriend,
    victim_physical = Victimofphysicalviolencee.g.assault,
    victim_property = Victimofpropertycrimee.g.thefthousebreaking,
    jailed_self = Detainedinjailcorrectionalfacility,
    jailed_family = Closefamilymemberdetainedinjailcorrectionalfacility,
    job_loss = Firedlaidofformaderedundantbyanemployer,
    financial_decline = Majorworseninginfinancialsituatione.g.wentbankrupt,
    natural_disaster = Aweatherrelateddisastere.g.floodbushfirecyclonedamagedordestroye,
    fire_or_explosion = Afireexplosiondamagedordestroyedyourhomeproperty,
    sexual_assault = Victimofsexualassaultviolence,
    transportation_accident = Serioustransportationaccidentcaraccidentboataccidentshipwreckair,
    other_life_event = Haveyouexperiencedanymajorlifeeventnotlistedaboveifyespleasespec
  )

# Verified!
names(life_event_data)

```

```{r summarize_life_events}

life_event_data %>%
  summarise(across(
    everything(),
    ~sum(. == 1, na.rm = TRUE)
  )) %>%
  pivot_longer(cols = everything(), names_to = "life_event", values_to = "count") %>%
  mutate(
    percent = round(100 * count / nrow(life_event_data), 1)
  ) %>%
  arrange(desc(count))

```

I have been curious to try the R package that can turn stuff into APA tables since that formatting is so annoying. So let's give it a go

```{r APA}

library(knitr)
library(kableExtra)

life_events_summary <- life_event_data %>%
  summarise(across(
    everything(),
    ~sum(. == 1, na.rm = TRUE)
  )) %>%
  pivot_longer(cols = everything(), names_to = "Life Event", values_to = "n") %>%
  mutate(
    Percent = round(100 * n / nrow(life_event_data), 1),
    `Life Event` = str_replace_all(`Life Event`, "_", " ") %>%
      str_to_title()
  ) %>%
  arrange(desc(n))

# Display as APA-style table
life_events_summary %>%
  kable(
    caption = "Frequencies and Percentages of Life Events Experienced",
    col.names = c("Life Event", "Frequency", "Percentage"),
    align = c("l", "c", "c"),
    format = "html"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center")

```

That is so cool! Not perfect, but much closer to APA format than a SPSS exported table. 

Now that these easier, preliminary analyses are replicated, I can move onto the other stuff. 